### 2021-05-03 initial/notes.tex
stash 05/03 bit more before meeting to do

The two key constraints we want are: (i) $\sim \frac{1}{5\omega_{\rm
1p}\epsilon}$ the tidal synchronization timescale (we will now use $t_a$ to
follow DL12 notation) should be faster than $\sim$ Gyr, (ii) $s_c$ ideally
should not be too small to forbid separatrix crossings. Note that $3 / Q' = 2k_2
/ Q$ ranges from around $10^{-2}$ for rocky Earths to $10^{-5}$ for Jupiters. We
can then write down the following dimensionless scalings (checked against the
dissipating disks paper w/ Dong):
% in paper 1, s_c = 2.04e4, for m_p = 0.01 Msun, a1=5AU, rho1=1g/cm^3, ap=30AU
\begin{align}
    \frac{s_c}{\Omega_1} &= \frac{
        \frac{3m_p}{4M_\star}\p{\frac{a_1}{\tilde{a}_{\rm p}}}^3
            }{\frac{3k_{\rm qp}}{2k_{\rm p}}
            \frac{M_\star}{M_1}\p{\frac{R_1}{a_1}}^3},\\
        &= \frac{m_{\rm p}m_1}{2M_{\star}^2}
            \p{\frac{k_{\rm p}}{k_{\rm qp}}}
            \p{\frac{a_1^6}{R_1^3 \tilde{a}_p^3}},\\
    \epsilon &=\frac{L}{2S}\frac{s}{2\Omega}\frac{1}{\omega_{\rm 1p}\cos I t_a}
            ,\\
    \frac{1}{t_a} &= \frac{9}{2Q'}\frac{M_\star}{m_{\rm 1}}
        \p{\frac{R_1}{a_1}}^5\Omega,\\
    \epsilon &= \frac{m_1a_1^2\Omega}{2(m_1R_1^2/4)s}\frac{s}{2\Omega}
        \frac{1}{\cos I}\p{\frac{9}{2Q'}\frac{M_\star}{m_1}
            \p{\frac{R_1}{a_1}}^5}
            \p{\frac{4M_\star a_p^3}{3m_pa_1^3}},\\
        &= \frac{1}{\cos I}
            \frac{6}{Q'}\p{\frac{M_\star^2}{m_1m_p}}
            \p{\frac{R_1^3a_p^3}{a_1^6}},\\
        &= \frac{1}{\cos I}\frac{3}{Q'}\frac{\Omega_1}{s_c},\\
% use paper I formula, change rho_1, mp, ap, a1
    \frac{s_c}{n_1} &= 0.06 \p{\frac{k_{\rm 1}}{k_{\rm q1}}}
            \p{\frac{m_{\rm p}}{0.001 M_{\odot}}}
            \p{\frac{a_{\rm p}}{5 \;\mathrm{AU}}}^{-3}
            \p{\frac{a_{\rm 1}}{0.3\;\mathrm{AU}}}^{6}
            \p{\frac{\rho_{\rm 1}}{3 \; \mathrm{g/cm}^3}}
            \p{\frac{M_\star}{M_{\odot}}}^{-2},\\
    \epsilon &\sim 10^{-2} \p{\frac{2k_2/Q}{10^{-3}}}.
\end{align}
Note that $2k_2/Q = 3/Q'$ in Dong's notation. The timescales now work since
$\omega_{\rm 1p} \sim 1.6\;\mathrm{Myr}$, so obliquities are driven to their
extreme values in around $0.8\;\mathrm{Gyr}$.

@@

>>>>

The two key constraints we want are: (i) $\sim \frac{1}{5\omega_{\rm
1p}\epsilon}$ the tidal synchronization timescale (we will now use $t_a$ to
follow DL12 notation) should be faster than $\sim$ Gyr, (ii) $s_c$ ideally
should not be too small to forbid separatrix crossings. Note that $3 / Q' = 2k_2
/ Q$ ranges from around $10^{-2}$ for rocky Earths to $10^{-5}$ for Jupiters. We
can then write down the following dimensionless scalings (checked against the
dissipating disks paper w/ Dong):
% in paper 1, s_c = 2.04e4, for m_p = 0.01 Msun, a1=5AU, rho1=1g/cm^3, ap=30AU
% NB: ta has masses flipped compared to Lai 2012, like Millholland 2020
\begin{align}
    \frac{s_c}{\Omega_1} &= \frac{
        \frac{3m_p}{4M_\star}\p{\frac{a_1}{\tilde{a}_{\rm p}}}^3
            }{\frac{3k_{\rm qp}}{2k_{\rm p}}
            \frac{M_\star}{M_1}\p{\frac{R_1}{a_1}}^3},\\
        &= \frac{m_{\rm p}m_1}{2M_{\star}^2}
            \p{\frac{k_{\rm p}}{k_{\rm qp}}}
            \p{\frac{a_1^6}{R_1^3 \tilde{a}_p^3}},\\
    \epsilon &=\frac{L}{2S}\frac{s}{2\Omega}\frac{1}{\omega_{\rm 1p}\cos I t_a}
            ,\\
    \frac{1}{t_a} &= \frac{9}{2Q'}\frac{M_\star}{m_{\rm 1}}
        \p{\frac{R_1}{a_1}}^5\Omega,\\
    \epsilon &= \frac{m_1a_1^2\Omega}{2(m_1R_1^2/4)s}\frac{s}{2\Omega}
        \frac{1}{\cos I}\p{\frac{9}{2Q'}\frac{M_\star}{m_1}
            \p{\frac{R_1}{a_1}}^5}
            \p{\frac{4M_\star a_p^3}{3m_pa_1^3}},\\
        &= \frac{1}{\cos I}
            \frac{6}{Q'}\p{\frac{M_\star^2}{m_1m_p}}
            \p{\frac{R_1^3a_p^3}{a_1^6}},\\
        &= \frac{1}{\cos I}\frac{3}{Q'}\frac{\Omega_1}{s_c},\\
% use paper I formula, change rho_1, mp, ap, a1
    \frac{s_c}{n_1} &= 0.06 \p{\frac{k_{\rm 1}}{k_{\rm q1}}}
            \p{\frac{m_{\rm p}}{0.001 M_{\odot}}}
            \p{\frac{a_{\rm p}}{5 \;\mathrm{AU}}}^{-3}
            \p{\frac{a_{\rm 1}}{0.3\;\mathrm{AU}}}^{6}
            \p{\frac{\rho_{\rm 1}}{3 \; \mathrm{g/cm}^3}}
            \p{\frac{M_\star}{M_{\odot}}}^{-2},\\
    \epsilon &\sim 10^{-2} \p{\frac{2k_2/Q}{10^{-3}}}.
\end{align}
Note that $2k_2/Q = 3/Q'$ in Dong's notation. The timescales now work since
$\omega_{\rm 1p} \sim 1.6\;\mathrm{Myr}$, so obliquities are driven to their
extreme values in around $0.8\;\mathrm{Gyr}$.

@@

\subsubsection{Physical Numbers}

We focus on the USP regime. In the notation of the paper, we need
\begin{align}
    1
        &\approx \frac{2n / \Omega_{\rm s} - \cos \theta
            }{\abs{g}t_{\rm a}\sin I}.
\end{align}
One tricky thing about this calculation is that $\eta$ depends on $\Omega_{\rm
s}$ which depends on $\cos \theta$ which depends on $\eta$:
\begin{align}
    \eta &= \eta_{\rm sync} \frac{n}{\Omega_{\rm s}},\\
    \frac{n}{\Omega_{\rm s}} &= \frac{1 + \cos^2\theta}{2 \cos \theta},\\
    \cos \theta &\approx \frac{\eta \cos I}{1 + \eta \sin I}.
\end{align}
This requires that
\begin{align}
    \frac{\eta}{\eta_{\rm sync}} &\approx \frac{1 + \eta \sin I
        }{2\eta \cos I},\\
    2\eta^2 \cos I &\approx \eta_{\rm sync} + \eta_{\rm sync}\eta \sin I,\\
    \eta &= \frac{\eta_{\rm sync}\sin I + \sqrt{
        \eta_{\rm sync}^2 \sin^2 I + 8\cos I \eta_{\rm sync}}}{
            4 \cos I}.
\end{align}
It proves sufficient to use the simpler form of this:
\begin{align}
    \eta &\approx \sqrt{\frac{\eta_{\rm sync}}{2\cos I}},\\
    t_{\rm a, c} &= \frac{2n / \Omega_{\rm s}}{\abs{g}\sin I},\\
        &= \frac{2}{\abs{g}\sin I\sqrt{2\eta_{\rm sync} \cos I}}.
\end{align}
Thus, we can continue to evaluate
\begin{align}
    1
        &\approx \frac{2\eta / \eta_{\rm sync} - \eta \cos I
            }{\abs{g}t_{\rm a}\sin I},\\
        &\approx \frac{1}{\sin I\sqrt{\eta_{\rm sync}\cos I}}
            0.003 \frac{1}{\cos I}\p{\frac{2 k_2 / Q}{10^{-3}}}
                \p{\frac{m_{\rm p}}{M_{\rm J}}}^{-1}
                \p{\frac{a_{\rm p}}{5\;\mathrm{AU}}}^{3}
                \p{\frac{a}{0.4\;\mathrm{AU}}}^{-6}
                \p{\frac{\rho}{3\;\mathrm{g/cm^3}}}^{-1}
                \p{\frac{M_{\star}}{M_{\odot}}}^2,\\
        &\approx \frac{0.005}{\sin I \cos^{3/2}I}
            \p{\frac{k}{k_q}}^{-1/2}
            \p{\frac{2 k_2 / Q}{10^{-3}}}
            \p{\frac{m_{\rm p}}{M_{\rm J}}}^{-3/2}
            \p{\frac{a_{\rm p}}{5\;\mathrm{AU}}}^{9/2}
            \p{\frac{a}{0.4\;\mathrm{AU}}}^{-9}
            \p{\frac{\rho}{3\;\mathrm{g/cm^3}}}^{-3/2}
            \p{\frac{M_{\star}}{M_{\odot}}}^3,\\
        &\approx \frac{3 \times 10^{-7}}{\sin I \cos^{3/2}I}
            \p{\frac{k}{k_q}}^{-1/2}
            \p{\frac{2 k_2 / Q}{10^{-3}}}
            % 1/5000
            \p{\frac{m_{\rm p}}{M_{\rm \oplus}}}^{-3/2}
            % 1e-9
            \p{\frac{a_{\rm p}}{0.05\;\mathrm{AU}}}^{9/2}
            % 1e9
            \p{\frac{a}{0.04\;\mathrm{AU}}}^{-9}
            % 1/3
            \p{\frac{\rho}{6\;\mathrm{g/cm^3}}}^{-3/2}
            \p{\frac{M_{\star}}{M_{\odot}}}^3,\\
    a_{\rm break}
        &\approx 0.01\;\mathrm{AU}
            \p{\frac{k}{k_q}}^{-1/18}
            \p{\frac{2 k_2 / Q}{10^{-3}}}^{1/9}
            \p{\frac{m_{\rm p}}{M_{\rm \oplus}}}^{-1/6}
            \p{\frac{a_{\rm p}}{0.05\;\mathrm{AU}}}^{1/2}
            \p{\frac{\rho}{6\;\mathrm{g/cm^3}}}^{-1/6}
            \p{\frac{M_{\star}}{M_{\odot}}}^{1/3}.
\end{align}
for $I = 5^\circ$. This seems a little bit on the low end.
